New approaches for boosting to uniformity

### 2014-10-02 assets/paperdraft/parts/metrics.tex
minor fixes in text

\def\knnSDE{\text{knnSDE}}

Though operating with bins is usually both simple and very efficient, 
in many cases it is hard to find the optimal size of bins in the space of uniform features (specifically in the case of more than two dimensions).
As mentioned earlier, problems can also arise due to bins with very low populations.
%One more situation when bins-based approach fails, is when we have too few events to obtain a good statistics at least in several bins.
In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself)
in the space of uniform features. Now we can compute the efficiency $\effknni$, from the empirical distribution $\Fknn$ of nearest neighbors. 
The weights for $\knni$ are proportional to the total weight of events in $\knni$:
\[
	\weightknni = \alpha \sum_{j \in \knni} w_j, \qquad \alpha^{-1} = \sum_i \sum_{j \in \knni} w_j,
\]
so again weights are normed to 1: $\sum_{i} \weightknni = 1$.

It is then possible to write the knn versions of SDE
\[
	\knnSDE^2(\globaleff)
		= \sum_{i \in \text{events}} \weightknni \abs{\effknni - \globaleff}^2,
\]
\[
	\knnSDE^2 = \frac{1}{k} \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]} \knnSDE^2(\globaleff),
\]
the Theil index of Efficiency
\[
	\text{knnTheil}(\globaleff) = \sum_{i \in \text{events}} \weightknni \; \frac{\effknni}{\globaleff} \; \ln{\frac{\effknni}{\globaleff}},
\]
\[
	\text{knnTheil} = \frac{1}{k} \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]} \text{knnTheil}(\globaleff),
\]
and the similarity-based measure:
\[
	 \sum_{i \in \text{events}} \weightknni \int \abs{\Fknn(x) - F(x)}^p dF(x).
\]

@@

\subsection{Advantages and Disadvantages of Different Metrics}
\subsubsection{Theil and SDE}
Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil. We have some linearly spaced array of mass of particles
from interval $[0,1]$, some constant $\alpha$ linearly spaced in interval $[.1,1]$. The predictions are correlated with mass via beta distribution.
We have two distributions that are different in a way that they are symmetrical. SDE should show no changes if we flip the distribution, while the Theil should make difference between pit and peak. First distribution with a peak is given by:

\[
	p\text{[i]}=\textbf{beta}(\alpha \cdot \bar m, \text{abs}(m\text[i]-\bar m))
\]

while the second one has minus sign in front. Figures~\ref{fig:peakdists}~and~\ref{fig:pitdists} show the distribution of the predictions
and classifier efficiency as a function of mass for the peak and pit distributions, respectively. 
\begin{figure}[h]
\centering
		\begin{subfigure}[b]{0.95\textwidth}
			\includegraphics[width=\textwidth]{graphs/PeakDistributions.png}
			\caption{Mass vs prediction.}
		\end{subfigure}
		\begin{subfigure}[b]{0.95\textwidth}
			\includegraphics[width=\textwidth]{graphs/PeakEffs.png}
			\caption{Mass vs efficiency. The different coloured lines correspond to
				different global classifier efficiencies, as explained in the legend on the leftmost subplot.}
		\end{subfigure}
		\caption{Peak response distribution and efficiencies as a function of mass. \label{fig:peakdists}}
\end{figure}

@@

Both metrics shown similar results, and that there is no significant difference on real data sets.
The SDE and Theil metrics can be used together in limit cases to more precisely determine nature of a distribution.

>>>>

\def\knnSDE{\text{knnSDE}}
\newcommand*\mean[1]{\overline{#1}}

Though operating with bins is usually both simple and very efficient, 
in many cases it is hard to find the optimal size of bins in the space of uniform features (specifically in the case of more than two dimensions).
As mentioned earlier, problems can also arise due to bins with very low populations.

In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself)
in the space of uniform features. Now we can compute the efficiency $\effknni$, from the empirical distribution $\Fknn$ of nearest neighbors. 
The weights for $\knni$ are proportional to the total weight of events in $\knni$:
\[
	\weightknni = \alpha \sum_{j \in \knni} w_j, \qquad \alpha^{-1} = \sum_i \sum_{j \in \knni} w_j,
\]
so again weights are normed to 1: $\sum_{i} \weightknni = 1$.

It is then possible to write the knn versions of SDE
\[
	\knnSDE^2(\globaleff)
		= \sum_{i \in \text{events}} \weightknni \abs{\effknni - \globaleff}^2,
\]
\[
	\knnSDE^2 = \frac{1}{k} \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]} \knnSDE^2(\globaleff),
\]
the Theil index of efficiency
\[
	\text{knnTheil}(\globaleff) = \sum_{i \in \text{events}} \weightknni \; \frac{\effknni}{\globaleff} \; \ln{\frac{\effknni}{\globaleff}},
\]
\[
	\text{knnTheil} = \frac{1}{k} \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]} \text{knnTheil}(\globaleff),
\]
and the similarity-based measure:
\[
	 \sum_{i \in \text{events}} \weightknni \int \abs{\Fknn(x) - F(x)}^p dF(x).
\]

@@

\subsection{Advantages and Disadvantages of Different Metrics}
\subsubsection{Theil and SDE}
Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil. We have some masses distributed uniformly in $[0,1]$, some constant $\alpha$ from interval $[.1,1]$. The predictions are correlated with mass via beta distribution.
We have two distributions that are different in a way that they are symmetrical. SDE should show no changes if we flip the distribution, while the Theil should make difference between pit and peak. 
First distribution with a peak is obtained by generating a prediction for each event according to it mass:
\[
	p \thicksim \text{Beta}(\alpha \mean{m}, \abs{m - \mean{m}})
\]
while the second one has minus sign in front. Figures~\ref{fig:peakdists}~and~\ref{fig:pitdists} show the distribution of the predictions
and classifier efficiency as a function of mass for the peak and pit distributions, respectively. 
\begin{figure}[h]
\centering
		\begin{subfigure}[b]{0.95\textwidth}
			\includegraphics[width=\textwidth]{graphs/PeakDistributions.png}
			\caption{Mass vs prediction.}
		\end{subfigure}
		\begin{subfigure}[b]{0.95\textwidth}
			\includegraphics[width=\textwidth]{graphs/PeakEffs.png}
			\caption{Mass vs efficiency. The different coloured lines correspond to
				different global classifier efficiencies, as explained in the legend on the leftmost subplot.}
		\end{subfigure}
		\caption{Peak response distribution and efficiencies as a function of mass. \label{fig:peakdists}}
\end{figure}

@@

Both metrics shown similar results, and that there is no significant difference on real data sets (as well as CvM).
The SDE and Theil metrics can be used together in limit cases to more precisely determine nature of a distribution.
